\MakeAbstract{
    在线教学与远程协作场景下，白板类应用需要同时应对高频输入导致的渲染与网络开销、弱网与断线重连下的一致性收敛，以及数学场景中低门槛的公式输入需求。本文面向“无需注册、快速创建/加入房间、短会话即时讨论”的使用方式，设计并实现轻量级实时数字协作白板系统 QuickBoard。系统采用三服务架构：前端基于 Vue~3 与 Fabric.js 实现对象级绘制与编辑；后端基于 Node.js（Express）与 Socket.IO 负责房间管理、实时事件分发与服务端权威裁决；OCR 服务基于 FastAPI 提供公式识别能力，并由后端统一代理返回 traceId 与错误码。同步一致性方面，系统实现“服务端权威裁决 + 字段级 LWW 合并 + roomVersion 增量补包 + 快照兜底 + 手动对齐”的组合策略，保证弱网与断线场景下的最终收敛。性能方面，通过视口裁剪与对象缓存降低大对象规模下的重绘开销；过程预览方面，通过 Ghost Brush 将高频预览与权威状态解耦，并采用批处理与限包降低消息风暴风险；公式能力方面，构建“框选—裁剪—识别—确认插入—可编辑渲染”的闭环。本文进一步给出一致性对照实验、渲染性能基准与降载测试的评估结果，为系统有效性提供可复核证据。
}{协作白板，实时同步，弱网鲁棒性，LWW，公式识别}

\MakeAbstractEng{
    Web-based whiteboards for online teaching and remote collaboration face three major engineering challenges: rendering and networking overhead under high-frequency interactions, eventual convergence under weak networks and reconnection, and low-barrier formula input in math-intensive scenarios. This thesis presents QuickBoard, a lightweight real-time collaborative whiteboard designed for no-registration, instant room creation/joining, and short-session discussions. The system adopts a three-service architecture: a Vue~3 + Fabric.js frontend for object-based drawing and editing, a Node.js (Express) + Socket.IO backend for room management, event distribution, and server-authoritative arbitration, and a FastAPI-based OCR service for formula recognition proxied by the backend with traceId and unified error codes. For synchronization, QuickBoard combines server-authoritative arbitration, field-level last-writer-wins (LWW) merging, version-based delta catch-up, snapshot fallback, and manual re-sync to achieve eventual convergence under weak networks. For performance, viewport culling and object caching reduce repaint costs at large object scales. For transient previews, Ghost Brush decouples preview traffic from the authoritative state and applies batching and per-message caps. For formulas, the system provides an end-to-end workflow from selection and recognition to confirmed insertion and editable rendering. The thesis further evaluates the system with reproducible evidence from a weak-network convergence experiment, a rendering benchmark, and a load-shedding test.
}{collaborative whiteboard, real-time synchronization, weak network robustness, LWW, mathematical OCR}
